---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AMPLE

<!-- badges: start -->
<!-- badges: end -->

AMPLE provides three Shiny apps to explore how Harvest Control Rules (HCRs) work and how performance indicators can be used to select between them.

The three apps are:

* *Introduction to HCRs* provides a simple overview to how HCRs work. Users are able to select their own HCR and step through its performance, year by year. Biological variability and estimation uncertainty are introduced.
* *Measuring performance* builds on the previous app and introduces the idea of using performance indicators to evaluate HCR performance.
* *Comparing performance* allows multiple HCRs to be created and tested, and their performance compared so that the preferred HCR can be selected.


## Installation

You can install the released version of AMPLE from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("AMPLE")
```

## Usage

Load the library as usual:

```{r eval=FALSE}
library(AMPLE)
```

The apps can be launched by calling the appropriate function:

```{r eval=FALSE}
intro_hcr()
```

```{r eval=FALSE}
measuring_performance()
```

```{r eval=FALSE}
comparing_performance()
```

The apps are also hosted online at:

* https://ofp-sam.shinyapps.io/AMPLE-intro-hcr/
* https://ofp-sam.shinyapps.io/AMPLE-measuring-performance/
* https://ofp-sam.shinyapps.io/AMPLE-comparing-performance/

## Vignettes

There are three vignettes that walk users through the app.
These can be used in workshops etc to help learning.

You can access them from R using:

```{r eval=FALSE}
vignette("intro_hcr", package="AMPLE")
vignette("measuring_performance", package="AMPLE")
vignette("comparing_performance", package="AMPLE")
```

The vignettes are also accessible through the Shiny apps, under the 'Information' tab.

## Deployment

To deploy the apps to shinyapps.io you need to make an 'app.R' file in the *root* directory of the package, i.e. in the same directory as the DESCRIPTION file.

The app.R file should have these contents:

```{r eval=FALSE}
# Include pkgload in DESCRIPTION import as we need to use it here.
pkgload::load_all(".")
intro_hcr() # Or whichever app function you want

# Include this stuff to make the publish button appear in RStudio - otherwise do rsconnect() by hand
#library(shiny)
#ui <- fluidPage(
#  "Hello, world!"
#  #br(),
#  #packageDescription("AMPLE")
#)
#server <- function(input, output, session) {
#}
#shinyApp(ui, server)
```

In RStudio, set the working directory to the package *root* directory.
Click publish and select all the files we need  (dropping man, vignettes, test folders).

Do not add this app.R file to the package when submitting to CRAN.
It should be listed in the .Rbuildignore.

