---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AMPLE

<!-- badges: start -->
<!-- badges: end -->

AMPLE provides three Shiny apps to explore how Harvest Control Rules (HCRs) work and how performance indicators can be used to select between them.

The three apps are:

* *Introduction to HCRs* provides a simple overview to how HCRs work. Users are able to select their own HCR and step through its performance, year by year. Biological variability and estimation uncertainty are introduced.
* *Measuring performance* builds on the previous app and introduces the idea of using performance indicators to evaluate HCR performance.
* *Comparing performance* allows multiple HCRs to be created and tested, and their performance compared so that the preferred HCR can be selected.


## Installation

You can install the released version of AMPLE from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("AMPLE")
```

## Usage

Load the library as usual:

```{r eval=FALSE}
library(AMPLE)
```

The apps can be launched by calling the appropriate function:

```{r eval=FALSE}
intro_hcr()
```

```{r eval=FALSE}
measuring_performance()
```

```{r eval=FALSE}
comparing_performance()
```

The apps are also hosted online at:

* URL 1
* URL 2
* URL 3

## Vignettes

There are three vignettes that walk users through the app.
These can be used in workshops etc to help learning.